---
title: "Weather Variables"
author: "Miles Rollison"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F)

library(tidyverse)
library(magrittr)
library(reticulate)

Sys.setenv(RETICULATE_PYTHON =  ".venv/bin/python")
```

```{r}
stations = tribble(~station, ~county, ~region,
                   'FTC03', 'Larimer', 'Northeast',
                   #'ALT01', 'Weld', 'Northeast',
                   'KSY01', 'Weld', 'Northeast',
                   #'FTL01', 'Weld', 'Northeast',
                   'EAC01', 'Adams', 'EastCentral',
                   'YUM02', 'Yuma', 'Northeast',
                   #'KRK01', 'Yuma', 'Northeast',
                   #'HXT01', 'Phillips', 'Northeast',
                   #'HYK02', 'Phillips', 'Northeast',
                   'AKR02', 'Washington', 'Northeast',
                   'BRL02', 'Kit Carson', 'EastCentral',
                   'LAM01', NA, 'Southeast',
                   'RFD01', NA, 'Southeast',
                   'HNE01', NA, 'Southeast',
                   'CTZ01', NA, 'Southwest',
                   'YJK01', NA, 'Southwest',
                   'TWC01', NA, 'Southwest',
                   'HOT02', NA, 'Mountain Area NW',
                   'FRT02', NA, 'Mountain Area NW',
                   'CTR01', NA, 'SLV',
                   'BLA01', NA, 'SLV'
                   )
write_csv(stations, paste0(fakepath, "thesis/Data/CoAgMet/station_county_info.csv"))
```

```{r pythonvars}
directory = paste0(fakepath, "thesis/Data/CoAgMet/")
stations = stations$station
```

```{python getdata}
import urllib.request
import os

os.chdir(r.directory)
stations = r.stations

####
#
#stations = ["FTC03", "ALT01","KSY02", "FTL01", "EAC01", "YUM02", "KRK01", "HXT01", "HYK02", "AKR02", "BRL02"]

####

variables = "station, date, mean_temp, max_temp, max_temp_time, min_temp, min_temp_time, vapor_press, max_rh, max_rh_time,min_rh, min_rh_time, solar_rad, wind_run, precip, max_soil_temp_5, max_soil_temp_5_time, min_soil_temp_5, min_soil_temp_5_time, max_soil_temp_15, max_soil_temp_15_time, min_soil_temp_15, min_soil_temp_15_time, voltage, year, max_wind, max_wind_time, max_wind_dir, et_asce, et_pk, sum_et_asce\n"

types = "variableTypes, 01-01-1001, 99.99, 99.99, 23:59:59, 99.99, 23:59:59, 99.99, 99.99, 23:59:59, 99.99, 23:59:59, 99.99, 99.99, 99.99, 99.99, 23:59:59, 99.99, 23:59:59,99.99, 23:59:59, 99.99, 23:59:59, 99.99, 1000L, 99.99, 23:59:59, 999.999, 99.99, 99.99, 99.99\n"

coagmet_before= "https://coagmet.colostate.edu/rawdata_results.php?start_year=1999&start_month=01&start_day=01&station="
#station_code = "STG01"
coagmet_after = "&end_year=2020&end_month=02&end_day=01&data_type%5B%5D=1&qc=1&etr=1"

#link = coagmet_before + station_code + coagmet_after

for station in stations:
    station_code = station 
    filename = station + '.csv'
    link = coagmet_before + station_code + coagmet_after

    f = urllib.request.urlopen(link)
    file = f.read()

    csv = str(file).replace('\\n', '\n')
    csv = variables + types + csv[2: len(csv)-1]
    outfile = open(filename, 'w')
    outfile.write(csv)
    outfile.close()
```


```{r verify}
tbl = read_csv(paste0(fakepath, "thesis/Data/CoAgMet/AKR02.csv"))
tbl
```

## Variables    
    Datetime (in the format YYYY-MM-DD)
    Mean Temperature (Celsius)
    Maximum Temperature (Celsius)
    Time of Maximum Temperature (HH:MM:SS)
    Minimum Temperature (Celsius)
    Time of Minimum Temperature (HH:MM:SS)
    Vapor Pressure (kPa)
    Maximum Relative Humidity (Fraction)
    Time of Maximum Relative Humidity (HH:MM:SS)
    Minimum Relative Humidity (Fraction)
    Time of Minimum Relative Humidity (HH:MM:SS)
    Solar Radiation (MJ/m^2)
    Wind_Run (km/day)
    Precipitation (millimeters)
    Maximum Soil Temperature at 5cm (Celsius)
    Time of Maximum Soil Temperature at 5cm (HH:MM:SS)
    Minimum Soil Temperature at 5cm (Celsius)
    Time of Minimum Soil Temperature at 5cm (HH:MM:SS)
    Maximum Soil Temperature at 15cm (Celsius)
    Time of Maximum Soil Temperature at 15cm (HH:MM:SS)
    Minimum Soil Temperature at 15cm (Celsius)
    Time of Minimum Soil Temperature at 15cm (HH:MM:SS)
    Battery Voltage
    Year of data
    Maximum daily wind gust (m/s)
    Time of maximum daily wind gust (min into day NOT HH:MM as other max/min)
    Direction of maximum daily wind gust (deg)
    ASCE Reference Evapotranspiration Model (mm)*
    Penman-Kimberly Ref. Evapotranspiration Model (mm)*
    Summed hourly ASCE Ref. ET (mm)*
