---
title: "Monthly Dataset Construction Code Cleanup"
author: "Miles Rollison"
date: "2/19/2020"
output: html_document
---

# Data Dictionary

price_lag:       bid_avg for previous observation of the same location, quality, and unit
price_diff:      difference between bid_avg and price_lag
date_diff:       months since last observation of the same location, quality, and unit
price_diff_adj:  price_diff standardized by dividing by date_diff
diff_pct:        price_diff in percentage terms
diff_pct_adj:    price_diff_adj in percentage terms
price_inc:       price increase or decrease 1 : increase; -1 : decrease; 0 : no change
price_dev:       diff_pct_adj measured in standard deviations
dev_cat:         price_dev as categorical variable with categories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F)

#library(reticulate)
#Sys.setenv(RETICULATE_PYTHON =  ".venv/bin/python")

library(tidyverse)
library(magrittr)
library(lubridate)

directory = function(x){paste0(fakepath, x)}
```

```{r functions}
make_lags = function(df, v, l= 1){
    for(i in seq(1:l)){
        df %<>% 
            mutate_at(v, function(x){lag(x, i)}) %>%
            rename_at(v, function(x){ paste0(x, "_lag", i) }) %>%
            full_join(df)   
    }
    return(df)
}
```

```{r hay, message=F}
st = Sys.time()
hay = read_csv(directory( "thesis/Data/AMS/OriginalReports/hay_reports_stacked.csv"))

hay %<>% separate(location, into = c("location", "state"), sep = ', ') %>% 
    filter(state == "CO") %>% 
    mutate(date = date(month),
           year = year(month),
           month = month(month),
           bid_avg = (bid_low + bid_high) / 2,
           spread = bid_high - bid_low,
           units = units%>% gsub(" Bales", '', .),
           quality = paste(class, `grade description`, sep = '_'),
           location = case_when(
               location == "Mountain Area" ~ "Mountains and Northwest",
               location == "Western Slope Area" ~ "Mountains and Northwest",
               TRUE ~ location)) %>% 
    rename(grade = `grade description`) %>% 
    select(year, month, everything(), -variety, -transmode, -`pricing point`, -`delivery period`) %>% 
    arrange(location, year, month) %>% 
    filter(quality != "ALFALFA_NA")

hay %<>% group_by(location, quality, units) %>%
    mutate(price_lag = lag(bid_avg),
           date_diff = as.numeric(date - lag(date)) %/% 28,
           price_diff = bid_avg - price_lag,
           price_diff_adj = ifelse(date_diff == 0, price_diff, price_diff / date_diff), 
           diff_pct = round(price_diff / bid_avg, 4) * 100,
           diff_pct_adj = round(price_diff_adj / bid_avg, 4) * 100,
           price_inc = sign(price_diff))

hay %<>% 
    mutate(price_dev = round(diff_pct_adj / (sd(diff_pct_adj, na.rm = T) + 0.0001)),
           dev_cat = case_when(price_dev > 3 ~ 4,
                               price_dev < -3 ~ -4,
                               TRUE ~ price_dev)) %>% 
    ungroup()

hay$price_inc = with(hay, factor(price_inc)) %>% as.ordered() 
hay$dev_cat = with(hay, factor(dev_cat)) %>% as.ordered() 


```

```{r corn, message=F}
corn = read_csv(directory( "thesis/Data/AMS/OriginalReports/corn_reports_stacked.csv"))

corn %<>% separate(location, into = c("location", "state"), sep = ', ') %>% 
    filter(state == "CO") %>% 
    mutate(date = date(month),
           year = year(month),
           month = month(month),
           bid_avg = (bid_low + bid_high) / 2,
           price_cwt = case_when(
               units == "Bushel" ~ (bid_avg / 54) * 100,
               TRUE ~ bid_avg),
           spread = bid_high - bid_low) %>% 
    select(date, year, month, everything(), -variety, -`grade description`, -`delivery period`, -class) %>% 
    filter(location != "Southwest Nebraska / Southeast Wyoming")

corn %<>% group_by(location) %>%
    mutate(price_lag = lag(bid_avg),
           price_diff = round(bid_avg - price_lag, 4),
           diff_pct = round(price_diff / bid_avg, 4),
           price_inc = sign(diff_pct)) %>% 
    ungroup()

corn$price_inc = with(corn, factor(price_inc)) %>% as.ordered() 

```

```{r drought, message=F}
dm = read_csv(directory( "thesis/Data/Drought Monitor/totalArea_categorical_byCounty_CO_20000101_20191219.csv"))   

counties = read_csv(directory( "thesis/Data/counties_in_regions_CO_FIPS.csv"))

#Aggregate to month level
dm %<>% group_by(year, month, County) %>%
    summarise_if(is.numeric, mean)

#Join with region classifications
dm %<>% ungroup %>% mutate(County = str_replace(County, " County", ''),
               year = as.integer(year),
               month = as.integer(month),
               D2plus = D2 + D3 + D4) %>% 
    full_join(counties) %>% 
    mutate_if(is.double, function(x) x / 1000)

dmh = dm %>% group_by(year, month, ams_hay_region) %>% 
    summarise_if(is.numeric, sum) %>% 
    select(-FIPS) %>% 
    arrange(year, month)

dmc = dm %>% group_by(year, month, ams_corn_region)  %>% 
    summarise_if(is.numeric, sum) %>% 
    select(-FIPS) %>% 
    arrange(year, month) 

dmh %<>% group_by(ams_hay_region) %>%
    mutate(drought_area_pct = 100 * (D2+D3+D4) / max(None),
           dry_area = 100 * (D0+D1+D2+D3+D4) / max(None)) %>% 
    make_lags( vars(D0, D1, D2, D3, D4, D2plus), 4)
dmc %<>% group_by(ams_corn_region) %>%
    mutate(dry_area = 100 * (D0+D1+D2+D3+D4) / max(None),
           drought_area_pct = 100 * (D2+D3+D4) / max(None)) %>% 
    make_lags( vars(D0, D1, D2, D3, D4, D2plus), 4)

dmh %<>% mutate(date = as.Date(paste(year, month, "01", sep = '-')),
    time_period = case_when(date <= as.Date("2005-04-30") ~ 1,
                                        date <= as.Date("2011-04-30") ~ 2,
                                        date <= as.Date("2014-10-01") ~ 3,
                                        date >  as.Date("2014-10-01") ~ 4)) %>% 
    select(-date)

dmc %<>% mutate(date = as.Date(paste(year, month, "01", sep = '-')),
    time_period = case_when(date <= as.Date("2005-04-30") ~ 1,
                                        date <= as.Date("2011-04-30") ~ 2,
                                        date <= as.Date("2014-10-01") ~ 3,
                                        date >  as.Date("2014-10-01") ~ 4)) %>% 
    select(-date)
```

```{r join}
hay %<>% full_join(dmh, by = c("location" = "ams_hay_region", "year", "month")) %>% arrange(year, month, location)
corn %<>% full_join(dmc, by = c("location" = "ams_corn_region", "year", "month")) %>% arrange(year, month, location)

rm(dm, dmc, dmh, counties)
#gc()
```

```{r write}
# write_csv(hay, directory( "thesis/Data/Constructed Datasets/hay_co_monthly_ams_dm.csv"))
# write_rds(hay, directory( "thesis/Data/Constructed Datasets/hay_co_monthly_ams_dm.Rds"))
# write_csv(corn, directory( "thesis/Data/Constructed Datasets/corn_co_monthly_ams_dm.csv"))
```

```{r}
spei = read_csv(directory("thesis/Data/Constructed Datasets/spei.csv"))


rename_locations = function(x){
    x %<>% mutate(location = case_when(
        location == "EC" ~ "East Central Colorado",
        location == "NE" ~ "Northeast Colorado",
        location == "NWMA" ~ "Mountains and Northwest",
        location == "SE" ~ "Southeast Colorado",
        location == "SLV" ~ "San Luis Valley",
        location == "SW" ~ "South Western Colorado",
        TRUE ~ "ERROR"))
    return(x)
}

spei = rename_locations(spei)
```

```{r}
timesteps = c(3, 4, 6, 12, 24)

for (i in timesteps) {
    df = spei %>% filter(TimeStep == i) %>% select(-id, -TimeStep)
    colnames(df) %<>% gsub("SPEI", paste0("SPEI", i), .)
    assign(paste0("spei", i), df, .GlobalEnv)
}


for (i in timesteps[2:5]) {
    name = paste0("spei", i)
    df = eval(as.name(name))
    spei3 %<>% full_join(df)
} 

rm(spei4, spei6, spei12, spei24, df)
#gc()
```

```{r}
hay %<>% full_join(spei3) %>% filter(!is.na(bid_avg))
corn %<>% full_join(spei3) %>% filter(!is.na(bid_avg))
```

```{r}
write_csv(hay, directory( "thesis/Data/Constructed Datasets/hay_co_monthly_ams_dm_spei.csv"))
write_rds(hay, directory( "thesis/Data/Constructed Datasets/hay_co_monthly_ams_dm_spei.Rds"))
write_csv(corn, directory( "thesis/Data/Constructed Datasets/corn_co_monthly_ams_dm_spei.csv"))
write_rds(corn, directory( "thesis/Data/Constructed Datasets/corn_co_monthly_ams_dm_spei.Rds"))
```

```{r}
rm(corn, hay, spei, spei3)
gc()

et = Sys.time()

et-st
```
